---
title: "NetworkGraphs"
author: "Catherine Laing"
date: "2023-06-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Infants' early words are phonologically similar to one another, if not in the vocabulary items they choose to produce, but in the way they produce them. This has been well-documented in a number of case studies [refs], and can be clearly observed in datasets of early productions. For example, Deuchar and Quay's [ref] record of their Spanish-English bilingual child's first [how many?] words shows that many of her first words are produced with an open CV syllable, and she produces a number of identical forms to refer to a range of different (though phonologically-similar) words. This suggests that infants may be drawing on a systematic approach to early productions, whereby a small subset of simple phonological forms are used to produce a range of more varied and phonologically-challenging adult targets. This paper will consider this systematicity from a networks perspective, drawing on network graphs of the early vocabulary to determine how similar infants' words are to one another, and how this changes over time.

In previous work [Laing, under review], I draw on network analysis to show that in the first three years of life, infants' acquisition and production can be predicted based on the words they already produce and how they produce them. That is, a word is more likely to be acquired if it is phonologically more similar to existing words in the productive repertoire, and if infants' production of the new word is more similar to their existing productions. This reflects the systematicity that has been well-documented in previous studies [refs]. Moreover, the phonological properties of infants' word productions are more similar to one another than their adult target forms would suggest. This seems obvious when we consider the reality of infants' early phonological productions, but, crucially, drawing on network analysis has shown to be effective in coding a large amount of early production data, which can be exceptionally time-consuming, and quantitatively analysing this data across a group of infants and a range of languages.

Network analysis is an increasingly popular method of analysing lexical acquisition. Network models allow the analysis of connectivity within a system (in our case, the lexical or phonological system), and can track how that connectivity changes over time. In the case of language development, the nodes (individual items) in the network typically consist of words, and these are connected (or not) depending on how similar two nodes are in phonological or semantic space. Because this is a convenient way to think about language development over time, a number of studies have considered vocabulary acquisition, in infants acquiring their first language [] and adults acquiring a novel language []. In the L1 acquisition literature, network analysis began with studies of semantic relatedness, testing the extent to which infants acquired clusters of words that were similar in semantic space, for example, animal words or labels for people [check refs]. More recently, work has started looking at the extent to which the words most commonly acquired in the early vocabulary, across a range of languages, are similar to one another on a phonological dimension. Contrary to the expectations set out here, these accounts did not present a clear case for phonological similarity in the developing vocabulary. However, as I set out in Laing [under review], this is likely because they draw on vocabulary norms from thousands of infants, whereby the individual variability in phonoloigcal acquisition is "averaged out". That is, where an individual infant may show a bias towards /b/-words in their early vocabulary, and another individual infant may show a bias towards /d/-words [see REF, who show that...], in vocabulary norm data the /b/-words and the /d/-words for these two infants are combined, meaning that any phonological similarity is diluted. Systematicity will thus be very difficult to identify when analysing data that does not allow for consideration of individual development trajectories. 

Another issue with these previous studies is the way in which phonological similarity is calibrated. As far as I am aware, all previous studies have drawn on Levenshtein (edit) distance when calculating the distance between one word in the network and another. By this measure, the phonetic nature of differences between forms is not taken into account, which is a particular issue when it comes to analysis of infant production. As an example, *pat*, *bat*, *cat*, *that* and *rat* all differ by an edit distance of 1, meaning that (by most measures) these forms would be equidistant from one another in a network. However, from a phonetic perspective the extent of the difference between these forms is highly variable. *Pat* and *bat* differ only in a voicing contrast, which in early production is rather redundant since voicing is not typically acquired until the second half of the second year [ref]; from an infant production perspective these forms may well be considered as one and the same. *Pat* and *cat* both have a voiceless consonant in onset position, and also share manner of articulation in onset position, but the place of articulation differs; these forms are similar, but not the same, phonetically speaking. On the other hand, *pat* is substantially more different than both *that* and *rat*, both in terms of the phonetic properties of the onset consonants (/ð/ and /r/ are both voiced, and differ in both place and manner of articulation from /p/). Crucially, while /p/, /b/ and /k/ are typically acquired very early on in development and even occur in the babble phase, both /ð/ and /r/ are acquired years later, not until age XXX by McLeod and Crowe's account [REF].

In Laing [under review] I addressed these issues by drawing on Euclidean distance as a measure of phonological distance. Following work by Monaghan and colleagues [ref], who used distinctive features from the government phonology school to take into account the more subtle phonetic differences between different speech sounds, this study calculated a 'distance score' between each word and each other word in the child's production (both actual productions and the adult target form) to create an individual network of lexical items for each infant in the dataset, that grew month-by-month as the vocabulary increased in size. This method has since been replicated by Kalinowski [ref], who also extended the approach to test the phonological properties of each word as individual 'feature bundles' [CHECK]. This approach differed from that of Laing [UR] because it did not compare speech sounds according to their phonological positioning within the word. Whereas Laing considered onset consonants in relation to other onset consonants, and codas in relation to other codas, etc, Kalinowski iterated across each word to establish an overall value of all the features of each word, regardless of what order the segments were produced in. This meant that if a child produced the form *shoe* as [uʃi] (with the consonant in S2 onset) and *sheep* as [ʃi] (with the consonant in S1 onset), the forms would still be considered as phonologically the same. This approach captures systematicity in early production even more appropriately, as it allows for the fact that often consonants may be similar across words (for example, through the presence of consonant harmony or reduplication), as well as between them. However, note that Kalinowski also tested the Euclidean distance method (and Levenshtein distance) in her data, and found consistent evidence for systematicity in the adult target words that Norwegian infants acquired over the first XX months of development. Note also that this analysis drew on vocabulary norming data, rather than naturalistic data.

Previous work, including that by Laing [UR], Kalinowski [in prep - CHECK] and others [refs], uses network growth algorithms to test whether learning can be predicted based on the words that infants already know or produce. If infants' early word acquisition and production is as systematic as previous work suggests, then we would expect that it is possible to predict learning based on the known words in the infant's existing lexicon. Methods and results differ between these accounts, as outlined in part above, but Laing, Kalinowski and Siew and Vitevitch all show that infants' and children's word learning can be predicted based on how likely new words are to connect to existing words in the lexicon (i.e. new words are more likely to be similar to known words). Laing [UR] compared this across adult target forms and infants' actual productions to find that infants' early words are phonologically more similar to one another than we would expect from analysing the target forms alone. That is, phonological systematicity may drive lexical selection, but drives production even more so.

These network growth models present a compelling computational approach to observing systematicity in the developing lexicon, but they don't give us a clear view into the extent to which early-acquired forms are produced in a phonologically similar way. That is, network growth models analyse connectivity (are two words similar, yes or no? if yes then they are connected in the network), rather than phonological distance (*how* similar are two words in the network?). This study expands on previous work, and builds on findings from Laing [UR] by analysing network graphs of infants' early lexicons. This provides an insight into the properties of the network, including how densely connected the network is, and the phonological distance between items in the network. Following Laing, this paper will analyse infants' actual productions as well as the adult target forms, to test the extent to which systematicity is present in prodution, in terms of both the words infants select in early development (Target forms), and how they produce them (Actual forms). 

- I probably want something about selected and adapted words

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
