---
title: "NetworkGraphs"
author:
- name: Catherine E. Laing
  affiliation: '1'
  corresponding: yes
  address: Department of Language and Linguistic Science, University of York, Heslington, YO10 5DD.
  email: catherine.laing@york.ac.uk
authornote: |
 All code and associated data for this manuscript can be found at https://github.com/cathelaing/NetworkGraphs.
shorttitle: Network graphs of early phonological development
output:
  papaja::apa6_pdf: default
  #extra_dependencies: ["float"]
  #papaja::apa6_word: default
abstract: "ADD."
keywords: systematicity, phonological development, preferential attachment, networks analysis
#wordcount: "" 
bibliography: Systematicity.bib
#floatsintext: yes
figsintext: no
figurelist: yes
tablelist: no
footnotelist: no
linenumbers: no
numbersections: false
mask: no
draft: no
documentclass: apa6
header-includes:
-  \DeclareDelayedFloatFlavor{kableExtra}{table}
- \usepackage{tipa}
#- \newfontfamily\PF{Arial}
classoption: man
affiliation:
- id: '1'
  institution: University of York, York, UK
latex_engine: xelatex
---

```{r setup, include=FALSE}

source("prelims.R")
r_refs(file="r-references.bib")


my_citations <- cite_r(
  file="r-references.bib"
  , pkgs=c("tidyverse", "igraph", "papaja", "lmerTest")
  , withhold=FALSE
  , footnote=TRUE
)

# load files

SWD_red <- globalsmallworlddata[which(complete.cases(globalsmallworlddata[
  ,c('path_length', 'clust_coef_global', 'clust_coef_avg', 'mean_k')])),]

stat_sum_df <- function(fun, geom="crossbar", ...) {
  stat_summary(fun.data=fun, colour="red", geom=geom, width=0.2, ...)
}

```

Infants' early words are phonologically similar to one another, if not in the vocabulary items they choose to produce, but in the way they produce them. This has been well-documented in a number of case studies [refs], and can be clearly observed in datasets of early productions. For example, Deuchar and Quay's [ref] record of their Spanish-English bilingual child's first [how many?] words shows that many of her first words are produced with an open CV syllable, and she produces a number of identical forms to refer to a range of different (though phonologically-similar) words. This suggests that infants may be drawing on a systematic approach to early productions, whereby a small subset of simple phonological forms are used to produce a range of more varied and phonologically-challenging adult targets. This paper will consider this systematicity from a networks perspective, drawing on network graphs of the early vocabulary to determine how similar infants' words are to one another, and how this changes over time.

Systematicity in early phonological development is most comprehensively discussed in work by Vihman [REFs]. In more than four decades of work incorporating data from hundreds of infants acquiring an impressive range of languages, Vihman demonstrates a clear systematicity in infants' path to target-like word production. From the initial 'surprisingly accurate' [ref?] forms that appear in the first stages of word production, infants are shown to draw on what they know: they generally choose, for first production, words that are simple in their target phonological form and that contain consonants that are already familiar from the most common syllables of canonical babble [@mccune_early_2001]. These forms are, in Vihman's terms, *selected* for first word production owing to their easily-producible features. As the vocabulary grows, infant must necessarily acquire forms that do not contain such accessible phonological or segmental properties. Here we begin to see regression in the accuracy of early production, as infants systematically adapt forms to fit the most common structures and segments in their repertoire. When words are systematically altered to fit an evidently dominant pattern in the child's output, these forms are said to be *adapted*. That is, systematicity is apparent not just in the earliest words, but across the trajectory of acquisition as infants deal with the challenges of early word production by relying on well-rehearsed output forms. That is, over the first months of lexical development at least, infants' productions of newly-acquired words are likely to match their productions of existing words in the lexicon.

In previous work [Laing, under review], I draw on network analysis to show that in the first three years of life, infants' acquisition and production can be predicted based on the words they already produce and how they produce them. That is, a word is more likely to be acquired if it is phonologically more similar to existing words in the productive repertoire, and if infants' production of the new word is more similar to their existing productions. This reflects the systematicity that has been well-documented in previous studies [refs]. Moreover, the phonological properties of infants' word productions are more similar to one another than their adult target forms would suggest. This seems obvious when we consider the reality of infants' early phonological productions, but, crucially, drawing on network analysis has shown to be effective in coding a large amount of early production data, which can be exceptionally time-consuming, and quantitatively analysing this data across a group of infants and a range of languages.

Network analysis is an increasingly popular method of analysing lexical acquisition. Network models allow the analysis of connectivity within a system (in our case, the lexical or phonological system), and can track how that connectivity changes over time. In the case of language development, the nodes (individual items) in the network typically consist of words, and these are connected (or not) depending on how similar two nodes are in phonological or semantic space. Because this is a convenient way to think about language development over time, a number of studies have considered vocabulary acquisition, in infants acquiring their first language [] and adults acquiring a novel language []. In the L1 acquisition literature, network analysis began with studies of semantic relatedness, testing the extent to which infants acquired clusters of words that were similar in semantic space, for example, animal words or labels for people [check refs]. More recently, work has started looking at the extent to which the words most commonly acquired in the early vocabulary, across a range of languages, are similar to one another on a phonological dimension. Contrary to the expectations set out here, these accounts did not present a clear case for phonological similarity in the developing vocabulary. However, as I set out in Laing [under review], this is likely because they draw on vocabulary norms from thousands of infants, whereby the individual variability in phonoloigcal acquisition is "averaged out". That is, where an individual infant may show a bias towards /b/-words in their early vocabulary, and another individual infant may show a bias towards /d/-words [see REF, who show that...], in vocabulary norm data the /b/-words and the /d/-words for these two infants are combined, meaning that any phonological similarity is diluted. Systematicity will thus be very difficult to identify when analysing data that does not allow for consideration of individual development trajectories. 

Another issue with these previous studies is the way in which phonological similarity is calibrated. As far as I am aware, all previous studies have drawn on Levenshtein (edit) distance when calculating the distance between one word in the network and another. By this measure, the phonetic nature of differences between forms is not taken into account, which is a particular issue when it comes to analysis of infant production. As an example, *pat*, *bat*, *cat*, *that* and *rat* all differ by an edit distance of 1, meaning that (by most measures) these forms would be equidistant from one another in a network. However, from a phonetic perspective the extent of the difference between these forms is highly variable. *Pat* and *bat* differ only in a voicing contrast, which in early production is rather redundant since voicing is not typically acquired until the second half of the second year [ref]; from an infant production perspective these forms may well be considered as one and the same. *Pat* and *cat* both have a voiceless consonant in onset position, and also share manner of articulation in onset position, but the place of articulation differs; these forms are similar, but not the same, phonetically speaking. On the other hand, *pat* is substantially more different than both *that* and *rat*, both in terms of the phonetic properties of the onset consonants (/ð/ and /r/ are both voiced, and differ in both place and manner of articulation from /p/). Crucially, while /p/, /b/ and /k/ are typically acquired very early on in development and even occur in the babble phase, both /ð/ and /r/ are acquired years later, not until age XXX by McLeod and Crowe's account [REF].

In Laing [under review] I addressed these issues by drawing on Euclidean distance as a measure of phonological distance. Following work by Monaghan and colleagues [ref], who used distinctive features from the government phonology school to take into account the more subtle phonetic differences between different speech sounds, this study calculated a 'distance score' between each word and each other word in the child's production (both actual productions and the adult target form) to create an individual network of lexical items for each infant in the dataset, that grew month-by-month as the vocabulary increased in size. This method has since been replicated by Kalinowski [ref], who also extended the approach to test the phonological properties of each word as individual 'feature bundles' [CHECK]. This approach differed from that of Laing [UR] because it did not compare speech sounds according to their phonological positioning within the word. Whereas Laing considered onset consonants in relation to other onset consonants, and codas in relation to other codas, etc, Kalinowski iterated across each word to establish an overall value of all the features of each word, regardless of what order the segments were produced in. This meant that if a child produced the form *shoe* as [uʃi] (with the consonant in S2 onset) and *sheep* as [ʃi] (with the consonant in S1 onset), the forms would still be considered as phonologically the same. This approach captures systematicity in early production even more appropriately, as it allows for the fact that often consonants may be similar across words (for example, through the presence of consonant harmony or reduplication), as well as between them. However, note that Kalinowski also tested the Euclidean distance method (and Levenshtein distance) in her data, and found consistent evidence for systematicity in the adult target words that Norwegian infants acquired over the first XX months of development. Note also that this analysis drew on vocabulary norming data, rather than naturalistic data.

Previous work, including that by Laing [UR], Kalinowski [in prep - CHECK] and others [refs], uses network growth algorithms to test whether learning can be predicted based on the words that infants already know or produce. If infants' early word acquisition and production is as systematic as previous work suggests, then we would expect that it is possible to predict learning based on the known words in the infant's existing lexicon. Methods and results differ between these accounts, as outlined in part above, but Laing, Kalinowski and Siew and Vitevitch all show that infants' and children's word learning can be predicted based on how likely new words are to connect to existing words in the lexicon (i.e. new words are more likely to be similar to known words). Laing [UR] compared this across adult target forms and infants' actual productions to find that infants' early words are phonologically more similar to one another than we would expect from analysing the target forms alone. That is, phonological systematicity may drive lexical selection, but drives production even more so.

These network growth models present a compelling computational approach to observing systematicity in the developing lexicon, but they don't give us a clear view into the extent to which early-acquired forms are produced in a phonologically similar way. That is, network growth models analyse connectivity (are two words similar, yes or no? if yes then they are connected in the network), rather than phonological distance (*how* similar are two words in the network?). This study expands on previous work, and builds on findings from Laing [UR] by analysing network graphs of infants' early lexicons. This provides an insight into the properties of the network, including how densely connected the network is, and the phonological distance between items in the network. It also allows us to make predictions about network properties over time that reflect word selection and adaptation. Following Laing [UR], this paper will analyse infants' actual productions as well as the adult target forms, to test the extent to which systematicity is present in production, in terms of both the words infants select in early development (Target forms), and how they produce them (Actual forms). 

Specifically, this paper addresses the following two questions:

1. How systematic are early word productions, and (how) does this change over time?
2. Are the phases of word selection and adaptation - characterised by target-like and non-target-like productions, respectively - identifiable in the dataset?

Network graphs will be generated using the *igraph()* package [@R-igraph] in R [@R-base]. To address the first question, properties of the graphs will be analysed to determine 1) how closely connected individual words are to one another; 2) how dense the overall distribution of words is in the network; and 3) how this changes over time. Following Vihman's work, and findings presented by Laing [UR], it is expected that the early vocabulary will become increasingly systematic over time. Following predictions made in previous work but not supported in the analyses, the analysis of network properties in this analysis is expected to show a decrease in systematicity later on in the data, as infants move towards target-like (and thus phonologically variable) production in the third year. To address the second question, network graphs of infants' actual productions will be compared with those of the target form, to trace the 'target-likeness' of individual productions, and how this changes over time. Following Vihman once again, early word selection would be reflected in early similarities between Actual and Target network properties, as target forms are selected to match the structures and segments that infants are able to produce. Over time, Actual and Target forms are expected to diverge, such that Actual forms show more systematicity in the data than Target forms. Again, by the third year, we may see the two come back together as novel words are produced in an increasingly target-like way. Variables used to test these predictions are outlined in detail below.


## Methods

# Data extraction and preparation

Data were drawn from two corpora on PhonBank (https://phonbank.talkbank.org/): Providence [American English, @demuth_word-minimality_2006] and Lyon [French, @demuth_katherine_prosodically-conditioned_2008]. These were selected due to their equivalent data collection methods and the fact that the infants' productions, as well as the corresponding target forms, are phonetically transcribed. Nine infants' (5 English, 4 French, 4 boys overall) data were extracted using Phon [@hedlund_gregory_phon_2020], from the transcript with their first-recorded word to the final transcript taken at age 2;6. Infants were recorded in the home on a fortnightly basis, participating in naturalistic interactions with their caregivers. Two of the American infants were recorded weekly during some periods of data collection, but this is not an issue for this analysis since no between-child comparisons will be made. See Demuth et al. [-@demuth_word-minimality_2006] and Demuth and Tremblay [-@demuth_katherine_prosodically-conditioned_2008] for full details of data collection.

Extracted data was filtered to include only words featuring on the communicative development inventory (CDI) of the respective language, including all variants of a given form. For example, plurals were included alongside the singular noun form, and variable verb conjugations alongside their corresponding infinitive verb form. These data were used to create a network for each infant, whereby all words in a given month that hadn't been produced in previous months were included as new items in the network. When a word had been produced in previous months, it was not included. While this means that the data does not capture change in the production of a single form over time, it allows us to observe network growth at the point of acquisition for each word form. 

To determine the structure of the network, the first step was to create distance values between each word and each other word in the network. This was first done using a 'global' network of all forms produced by the infant up to the final session at 2;6, to create a large distance matrix for each infant that incorporated all word productions. Distance values were generated using methods set out in Monaghan et al. [-@monaghan_measures_2010], using distinctive features to generate a set of phonetic values for each word that could then be compared with all other words (note that only consonants were analysed, given that vowels are highly variable in early production and also very difficult to transcribe accurately; [@donegan_normal_2013; @kent_what_2020]). Euclidean distance between the values of each word and each other word in each infant's global network was then used to determine how close/distant words were from one another. Often, infants produced multiple tokens of the same word type in a given month, often with high variability across tokens. Because it was not possible to generate networks with all tokens included (even with only single word types included, the final dataset for all nine infants includes over 5.5 million data points), a mean value for each distinctive feature was established across tokens, meaning that each word's distinctive feature value represents the variability of the infant's production of a given word. This may not be a perfect measure, but it is more representative than taking, for example, the first instance of each word type.

Distance scores were generated between each word and each other word in each child's dataset, for both Target and Actual forms. These scores were then normalised, and a normalised distance of 0.25 was chosen to indicate connectivity. That is, words were said to be connected in the network if their distance score was ≤0.25. This accounted for the lower quartile of connectivity across the dataset. See Supplementary Materials for an overview of how this measure was established.

adapt this bit to fit with the data used in this analysis (and re-write):

<!-- The final dataset includes `r round(types$n)` word types overall, aggregated across infants (English=`r round(subset(types.corpus, corpus == "English")$n)`, French=`r round(subset(types.corpus, corpus == "French")$n)`). On average, there were `r round(tokens$mean_tok)` tokens of each word type (SD = `r round(tokens$sd_tok)`). See Table \@ref(tab:table-data-overview) for a breakdown by corpus and child. All but `r syl.4` tokens (all French) in the data had three syllables or fewer in the target form, with `r round(subset(syls, corpus == "English")$mean_syl)` syllable on average in the English data (SD = `r round(subset(syls, corpus == "English")$sd_syl,2)`) and `r round(subset(syls, corpus == "French")$mean_syl,2)` in the French data (SD = `r round(subset(syls, corpus == "French")$sd_syl,2)`). -->

# Network graphs

The prepared data was then used to generate a series of network graphs for each infant (for both Target and Actual data) using the *igraph()* package in R [@R-igraph]. One network was generated per month, for each month in the dataset, and these grew cumulatively such that the network at time-point *n*+1 included all words at time-point *n*, plus all additional words produced for the first time at *n*+1.  

The *igraph()* package generates graphs that include all nodes (whether or not they are connected to other nodes), and accounts for the distance between nodes, as well as the clustering of nodes in graphical space. 

Two key variables will be explored through an analysis of network graphs: *mean path length* and *clustering coefficient*. Path length is a measure of distance between nodes, and mean path length indexes the average connectivity (of all connected nodes) within a network; by this measure, we would expect that systematicity in early phonological development would be reflected in low mean path length. Clustering coefficient is an indication of network density: a higher density of nodes in the network indicates denser clusters of similar forms; again, this is what we would expect to see in a network of early phonological development. See Goldbeck [-@golbeck_analyzing_2013] for a full overview of network structures and measures.



## Results

1. How systematic are early word productions, and (how) does this change over time?

```{r GAMM prep, echo=F, message=FALSE, warning=FALSE, comment=F}

SWD_actual <- subset(SWD_red, data_type == "actual")
SWD_target <- subset(SWD_red, data_type == "target")

SWD_actual$start.event <- SWD_actual$session_ordinal == 1 
SWD_target$start.event <- SWD_target$session_ordinal == 1 

SWD_red$start.event <- SWD_red$session_ordinal == 1 


```

```{r GAMM age actual MPL, message=FALSE, warning=FALSE, include=FALSE}

MPL.gamm.base_A <- bam(path_length ~ 
                       corpus +
                       s(age, bs = "cr") +                       
                       s(age, by=Speaker, bs="cr") +
                       s(age, by=corpus, bs="cr"),
                     dat=SWD_actual, method="ML")

rA <- start_value_rho(MPL.gamm.base_A) 

MPL.gamm.1_A <- bam(path_length ~ 
                       corpus +
                      s(age, bs = "cr") + 
                      s(age, corpus, bs="fs", m=1, k=2) +
                      s(age, Speaker, bs="fs", m=1, k=9),  
                        data = SWD_actual, method = "ML", 
                        rho=rA, AR.start=SWD_actual$start.event)

MPL.gamm.0_A <- bam(path_length ~ 
                       corpus +
                      #s(age, bs = "cr") + 
                      s(age, corpus, bs="fs", m=1, k=2) +
                      s(age, Speaker, bs="fs", m=1, k=9),  
                        data = SWD_actual, method = "ML", 
                        rho=rA, AR.start=SWD_actual$start.event)

MPL.gamm.actual <- compareML(MPL.gamm.1_A, MPL.gamm.0_A)
MPL.gamm.A <- MPL.gamm.actual$table

MPL.gamm.A_summ <- summary(MPL.gamm.A)

```

```{r GAMM age actual CC, message=FALSE, warning=FALSE, include=FALSE}

CC.gamm.base_A <- bam(clust_coef_avg ~ 
                       corpus +
                       s(age, bs = "cr") +                       
                       s(age, by=Speaker, bs="cr") +
                       s(age, by=corpus, bs="cr"),
                     dat=SWD_actual, method="ML")

rA <- start_value_rho(CC.gamm.base_A) 

CC.gamm.1_A <- bam(clust_coef_avg ~ 
                       corpus +
                      s(age, bs = "cr") + 
                      s(age, corpus, bs="fs", m=1, k=2) +
                      s(age, Speaker, bs="fs", m=1, k=9),  
                        data = SWD_actual, method = "ML", 
                        rho=rA, AR.start=SWD_actual$start.event)

CC.gamm.0_A <- bam(clust_coef_avg ~ 
                       corpus +
                      #s(age, bs = "cr") + 
                      s(age, corpus, bs="fs", m=1, k=2) +
                      s(age, Speaker, bs="fs", m=1, k=9),  
                        data = SWD_actual, method = "ML", 
                        rho=rA, AR.start=SWD_actual$start.event)

CC.gamm.actual <- compareML(CC.gamm.1_A, CC.gamm.0_A)
CC.gamm.A <- CC.gamm.actual$table

CC.gamm.A_summ <- summary(CC.gamm.A)

```

```{r GAMM data type actual MPL, message=FALSE, warning=FALSE, include=FALSE}

MPL.gamm.base_DT <- bam(path_length ~ 
                         data_type +
                       corpus +
                       s(age, bs = "cr") +                       
                       s(age, by=Speaker, bs="cr") +
                         s(age, by=data_type, bs="cr") +
                       s(age, by=corpus, bs="cr"),
                     dat=SWD_red, method="ML")

rDT <- start_value_rho(MPL.gamm.base_DT) 

MPL.gamm.1_DT <- bam(path_length ~ 
                      data_type +
                       corpus +
                      s(age, bs = "cr") + 
                      s(age, corpus, bs="fs", m=1, k=2) +
                      s(age, data_type, bs="fs", m=1, k=2) +
                      s(age, Speaker, bs="fs", m=1, k=9),  
                        data = SWD_red, method = "ML", 
                        rho=rDT, AR.start=SWD_red$start.event)

MPL.gamm.0_DT <- bam(path_length ~ 
                      #data_type +
                       corpus +
                      s(age, bs = "cr") + 
                      s(age, corpus, bs="fs", m=1, k=2) +
                      #s(age, data_type, bs="fs", m=1, k=2) +
                      s(age, Speaker, bs="fs", m=1, k=9),  
                        data = SWD_red, method = "ML", 
                        rho=rDT, AR.start=SWD_red$start.event)

MPL.gamm.DT <- compareML(MPL.gamm.1_DT, MPL.gamm.0_DT)
MPL.gamm.DT.prep <- MPL.gamm.DT$table

MPL.gamm.DT_summ <- summary(MPL.gamm.DT.prep)


plot_diff(MPL.gamm.1_DT, view="age", comp=list(data_type=c("target","actual")),
  # main = "Figure 3",
  # ylab = "Est. difference in scaled PAT values",
  xlab = "Age (months)",
  xlim = c(10,30),
  hide.label = TRUE)
```

```{r network graph model comparisons, message=FALSE, warning=FALSE, include=FALSE}


# Mean path length

global_pathlength_model <- lmerTest::lmer(path_length ~ data_type + age + corpus + (1+age|Speaker),
                                             data=subset(SWD_red,
                                                           data_type %in% c("target", "actual")
                                                           ),
                                       REML=FALSE)
#summary(global_pathlength_model)

global_pathlength_model_null <- lmerTest::lmer(path_length ~ age + corpus + (1+age|Speaker),
                                             data=subset(SWD_red,
                                                           data_type %in% c("target", "actual")
                                                           ),
                                       REML=FALSE)

pathlength <- anova(global_pathlength_model, global_pathlength_model_null)

pathlength_df <- pathlength$Df[2]
pathlength_chisq <- pathlength$Chisq[2]
pathlength_p.value <- pathlength$`Pr(>Chisq)`[2]

# Clustering coefficient

global_clust_model <- lmerTest::lmer(clust_coef_avg ~ data_type + age + corpus + (1+age|Speaker),
                                             data=subset(SWD_red,
                                                           data_type %in% c("target", "actual")
                                                           ),
                                       REML=FALSE)
#summary(global_clust_model)

global_clust_model_null <- lmerTest::lmer(clust_coef_avg ~ age + corpus + (1+age|Speaker),
                                             data=subset(SWD_red,
                                                           data_type %in% c("target", "actual")
                                                           ),
                                       REML=FALSE)

clustcoef <- anova(global_clust_model, global_clust_model_null)

clustcoef_df <- clustcoef$Df[2]
clustcoef_chisq <- clustcoef$Chisq[2]
clustcoef_p.value <- clustcoef$`Pr(>Chisq)`[2]

```

```{r network graphs model outputs, include=FALSE}

table.model.output <- rbind(pathlength, clustcoef) %>%
  rownames_to_column(var="Model") %>%
  filter(Chisq > 0) %>%
  rename(#"PA Df"="Df",
         #"PA Chisq"="Chisq",
         "p"=`Pr(>Chisq)`) %>%
  mutate(Variable = ifelse(Model == "global_pathlength_model", "Path Length", "Clustering Coefficient"),
             p=scales::pvalue(p)) %>%
  select(Variable, `Df`, `Chisq`, `p`)

pathlength.summary <- summary(global_pathlength_model)

pathlength.summary_tab <- pathlength.summary$coefficients %>%
  as.data.frame %>%
  rename(
    "beta"="Estimate"
     , "SE"="Std. Error"
     , "t"="t value"
     , "p"="Pr(>|t|)") %>%
  mutate(
    Effect=papaja:::beautify_terms(rownames(.))) %>%
  printnum(
    digits=c(2, 2, 2, 3)
    , gt1=c(TRUE, TRUE, TRUE, TRUE)
    , zero=c(TRUE, TRUE, TRUE, TRUE)
  ) %>%
  mutate(beta=as.numeric(beta),
        SE=as.numeric(SE),
        `t`=as.numeric(`t`),
        `p`=as.numeric(`p`),
             p=scales::pvalue(p)) %>%
  mutate(Effect = fct_recode(Effect,
                             "Data type" = "Data typetarget",
                             "Corpus" = "CorpusFrench")) %>%
  dplyr::select(Effect, beta, SE, t, p)

clustcoef.summary <- summary(global_clust_model)

clustcoef.summary_tab <- clustcoef.summary$coefficients %>%
  as.data.frame %>%
  rename(
    "beta"="Estimate"
     , "SE"="Std. Error"
     , "t"="t value"
     , "p"="Pr(>|t|)") %>%
  mutate(
    Effect=papaja:::beautify_terms(rownames(.))) %>%
  printnum(
    digits=c(2, 2, 2, 3)
    , gt1=c(TRUE, TRUE, TRUE, TRUE)
    , zero=c(TRUE, TRUE, TRUE, TRUE)
  ) %>%
  mutate(beta=as.numeric(beta),
        SE=as.numeric(SE),
        `t`=as.numeric(`t`),
        `p`=as.numeric(`p`),
             p=scales::pvalue(p)) %>%
  dplyr::select(beta, SE, t, p)

network.model.summary <- cbind(pathlength.summary_tab, clustcoef.summary_tab)

rownames(network.model.summary) <- NULL

PL_beta <- (subset(pathlength.summary_tab, Effect == "Data type")$beta)*100

```

Mean path length and clustering coefficient were calculated for each child's monthly networks. Target and Actual data was then compared using linear mixed-effects regression. Mean path length and clustering coefficient were included as dependent variables in two individual models, with Data type (Actual vs. Target), age and corpus as fixed effects and subject as a random effect with a by-subject random slope for the effect of age. Nested model comparisons showed that data type significantly predicted the mean path length within a network ($\chi^2 (`r pathlength_df`)$=`r pathlength_chisq`, *p*`r printp(pathlength_p.value)`) but not clustering coefficient ($\chi^2 (`r clustcoef_df`)$=`r clustcoef_chisq`, *p*=`r printp(clustcoef_p.value)`); mean path length was  `r PL_beta`% higher in Target compared with Actual data. See Table \@ref(tab:table-network-model-outputs) and Figures \@ref(fig:figure-path-length) and \@ref(fig:figure-clust-coef).

```{r table-network-model-outputs, echo=F, message=FALSE, warning=FALSE, comment=F, results="asis"}
cap="Outputs from linear mixed effects regression models testing the effect of Data type (Actual vs. Target, Actual as baseline) on mean path length and clustering coefficient as dependent variables, with corpus (English vs. French) and Age as additional fixed effects and subject as a random effect. By-subject random slopes were included for the effect of Age."
knitr::kable(network.model.summary, "latex", booktabs=T, longtable=T, caption=cap, digits=2, align="c") %>%
  add_header_above(c(" "=1, "Mean Path Length"=4, "Clustering Coefficient"=4)) %>%
  kable_styling()
```


```{r figure-path-length, echo=FALSE, fig.cap=cap, fig.pos='H', message=FALSE, warning=FALSE}
plot(path_length_plot)
cap <- sprintf("Change in mean path length over time, in Actual compared with Target data. Coloured lines represent Data type; coloured bands represent 95%% CIs.")
```

```{r figure-clust-coef, echo=FALSE, fig.cap=cap, fig.pos='H', message=FALSE, warning=FALSE}
plot(clust_coef_plot)
cap <- sprintf("Change in average clustering coefficient over time, in Actual compared with Target data. Coloured lines represent Data type; coloured bands represent 95%% CIs.")
```title: "OLD"
output: pdf_document
date: "2023-03-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


 



You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
